The contribution of blowing snow to cloud properties and the atmospheric radiative budget over Antarctica

### 2020-10-27 manuscript/manuscript.tex
WIP: paper

\author
{Stefan Hofer,$^{1\ast}$ Charles Amory,$^{2}$ Christoph Kittel,$^{2}$ and Trude Storelvmo$^{1}$ \\
\\
\normalsize{$^{1}$Department of Geosciences, University of Oslo, Oslo, Norway}\\
\normalsize{$^{2}$Laboratory of Climatology, Department of Geography, University of Liège, Belgium}\\
\\
\normalsize{$^\ast$Corresponding author: Stefan Hofer, stefan.hofer@geo.uio.no}
}

@@

Explicitly modelling blowing in snow in MAR leads to a notable change in the atmospheric structure of the lowermost 100s of meters above ground (Fig.\ref{fig:Test} A-C). Over the flat interior of the Antarctic Ice Sheet, the first few 100 m show a strong decrease in atmospheric temperature, with a mean 0-500 m difference of \textbf{XX $\pm$ XX \textdegree C} in elevations greater than 2000 m above mean sea level (Fig.\ref{fig:Test} A). This cooling of the near-surface atmosphere in our MAR blowing snow simulations, when compared to MAR without blowing snow, is most likely due to the efficient sublimation of airborne blowing snow particles due to the dry (and cold) surroundings over the interior Antarctic plateau. Conversely, over the steep slopes and the flat, low-lying ice shelves surrounding the grounded ice, this decrease in temperature in the blowing snow simulations is less notable. The mean 0-500 m above surface difference lies at \textbf{XX pm XX \textdegree C}. The contrasting picture between the flat interior and the steeper and lower margins of Antarctica is likely due to two effects: I) Due to the on average lower relative humidity and the greater water vapour gradients, the airborne blowing snow particles are more readily sublimated in the drier interior plateau than over the margins. This change of phase from solid to gaseous requires energy from the surrounding air to break up the bonds between the H$_2$O molecules (\textbf{CITE CC}), leading to a drop in temperature. II) Due to strong adiabatic mixing and turbulence in areas where the gravitational pull accelerates the katabatic winds over steep terrain, the snow particles are not confined to the very stable boundary layer anymore. Therefore, the sublimational cooling is less concentrated and subsequently also lower in magnitude over the margins of Antarctica and the ice shelves.

@@

>>>>

\author
{Stefan Hofer,$^{1\ast}$ Charles Amory,$^{2}$ Christoph Kittel,$^{2}$ Tim Carlsen$^{1}$ and Trude Storelvmo$^{1}$ \\
\\
\normalsize{$^{1}$Department of Geosciences, University of Oslo, Oslo, Norway}\\
\normalsize{$^{2}$Laboratory of Climatology, Department of Geography, University of Liège, Belgium}\\
\\
\normalsize{$^\ast$Corresponding author: Stefan Hofer, stefan.hofer@geo.uio.no}
}

@@

Explicitly modelling blowing in snow in MAR leads to a notable change in the atmospheric structure of the lowermost 100s of meters above ground (Fig.\ref{fig:Test} A-C). Over the flat interior of the Antarctic Ice Sheet, the first few 100 m show a strong decrease in atmospheric temperature, with a mean 0-500 m difference of \textbf{XX $\pm$ XX \textdegree C} in elevations greater than 2000 m above mean sea level (Fig.\ref{fig:Test} A). This cooling of the near-surface atmosphere in our MAR blowing snow simulations, when compared to MAR without blowing snow, is most likely due to the efficient sublimation of airborne blowing snow particles due to the dry (and cold) surroundings over the interior Antarctic plateau. Conversely, over the steep slopes and the flat, low-lying ice shelves surrounding the grounded ice, this decrease in temperature in the blowing snow simulations is less notable. The mean 0-500 m above surface difference lies at \textbf{XX pm XX \textdegree C}. The contrasting picture between the flat interior and the steeper and lower margins of Antarctica is likely due to two effects: 1) Due to the on average lower relative humidity and the greater water vapour gradients, the airborne blowing snow particles are more readily sublimated in the drier interior plateau than over the margins. This change of phase from solid to gaseous requires energy from the surrounding air to break up the bonds between the H$_2$O molecules (\textbf{CITE CC}), leading to a drop in temperature. 2) Due to strong adiabatic mixing and turbulence in areas where the gravitational pull accelerates the katabatic winds down steep terrain, the snow particles are not confined to the very stable boundary layer anymore. Therefore, the sublimational cooling is less concentrated and subsequently also lower in magnitude over the margins of Antarctica and the ice shelves.

@@

In the boundary layer, accounting for drifting snow also increases cloud formation over the Antarctic continent (Fig.\ref{fig:Test}B). Our results show that the strongest increase in 2000-2019 average cloud cover over the interior plateau strongly overlap with the changes in temperature seen in Figure \ref{fig:Test}A. In elevations above 2000 m above mean sea level the lowermost 500 m of the atmosphere show an increase of \textbf{XX $\pm$ XX\%} in cloud cover. However, there are three overlapping mechanisms that can explain the greater cloud amount over the Antarctica, when accounting for blowing snow. 1) Blowing snow particles can act as additional nuclei on which ice can grow or help with ice growth through the Wegener-Bergeron-Findeisen process. Ice crystal number concentration can potentially multiplied through secondary ice processes \textbf{cite Sotiropoulou}. 2) The sublimation of airborne snow particles leads to a cooling of the surrounding air, while increasing the specific humidity, both bringing the environment closer to saturation (\textbf{Cite Amory saturation}). 3) Thick blowing snow layers themselves act as a cloud, due to their ability to interact with incoming solar radiation (i.e. a cloud optical depth > 0) and their influence on the atmospheric longwave emissivity (i.e. they increase the atmospheric longwave emissivity $\epsilon$). It is likely that in most cases these three processes can act simultaneously. Additionally, even state-of-the-art active satellite products like the \textbf{Cloudsat-Calipso XX}, have a vertical resolution of 240-480 m when looking at their cloud-related datasets. This limitation in vertical resolution renders it quite unlikely that blowing snow is detected as a cloud from remote sensing platforms, especially over highly-reflective surfaces.

Accounting for blowing snow also alters the cloud radiative effect (Fig.\ref{fig:Test} C). While we see the strongest effects again in the boundary layer, especially over the steeper margins the CRE is altered to elevations of up to 5000 m. This vertical influence on the CRE might be due to the fact that airborne blowing snow particles can be mixed to layers above the boundary layer in zones with stronger adiabatic mixing and turbulence, i.e. over the steeper slopes where the katabatic winds are the strongest. Subsequently, these additional ice crystals can influence the macrophysical cloud properties (IWP, LWP and COD), and therefore the CRE. Additionally, because of changes in the vertical temperature distribution and humidity due to blowing snow sublimation, also the effectiveness and temperature of the layers that emit the longwave radiation can be altered between the two simulations.
